---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Travis build status](https://travis-ci.com/poissonconsulting/dttr2.svg?branch=master)](https://travis-ci.com/poissonconsulting/dttr2)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/poissonconsulting/dttr2?branch=master&svg=true)](https://ci.appveyor.com/project/poissonconsulting/dttr2)
[![Coverage status](https://codecov.io/gh/poissonconsulting/dttr2/branch/master/graph/badge.svg)](https://codecov.io/github/poissonconsulting/dttr2?branch=master)
[![License: MIT](https://img.shields.io/badge/License-MIT-green.svg)](https://opensource.org/licenses/MIT)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/dttr2)](https://cran.r-project.org/package=dttr2)

# dttr2

`dttr2` (Dates, Times and a TimeR) is an R package to manipulate date (Date), date time (POSIXct) and time (hms) vectors.
It provides functions to perform simple conversions and manipulations.
It also provides a timer.

Key design principles include

1. use existing classes
2. date times are discrete
3. floor date times
4. wrap times
5. keep the same time zone
6. use a default origin

## Installation

To install the latest development version from the Poisson drat [repository](https://github.com/poissonconsulting/drat)
```
if(!"drat" %in% installed.packages()[,1]) 
  install.packages("drat")
drat::addRepo("poissonconsulting")
install.packages("dttr2")
```

To install the latest development version from [GitHub](https://github.com/poissonconsulting/checkr)
```
if(!"devtools" %in% installed.packages()[,1]) 
  install.packages("devtools")
devtools::install_github("poissonconsulting/err")
devtools::install_github("poissonconsulting/checkr")
devtools::install_github("poissonconsulting/dttr2")
```

## Introduction

### Time Units

`dttr2` floors any objects it encounters to the nearest time unit (days for Date and seconds for POSIXct and hms).

```{r}
library(dttr2)
date <- as.Date(1.75, origin = "1970-01-01")
date
as.numeric(date)

date <- dtt_date(date)
date
as.numeric(date)
```

Similarly, the time unit accessor (`dtt_second()`, `dtt_minute()` etc) and settor functions as well as the functions to add (ie `dtt_add_seconds()`) and subtract time units return or require integers.

```{r}
dtt_day(date)
dtt_add_days(date, 2L)
dtt_add_months(date, 24L)
```

### Times

`dttr2` wraps negative times or times greater than 23:59:59 to the actual clock time.

```{r}
time <- hms::as_hms(-3)
time
dtt_time(time)
dtt_time(hms::as_hms("24:00:00"))
```

### Time Zones

The user can override the default time zone (UTC) when handling POSIXct vectors using `dtt_set_default_tz()`.

```{r}
dtt_default_tz()
date_time <- dtt_date_time("1970-01-01 03:00:00")
dtt_set_default_tz("Etc/GMT+10")
dtt_default_tz()
date_time
```

They can assign a new time zone to a POSIXct object whilst leaving the clock time unchanged using `dtt_set_tz()` or adjust the time zone so that clock (but not the actual) time is altered using `dtt_adjust_tz()` 

```{r}
date_time
dtt_set_tz(date_time)
dtt_adjust_tz(date_time)
```

### Combining Dates and Times

`dttr2` makes it easy to decompose POSIXct vectors into Date and hms vectors and then recombine them again.

As neither dates nor times have associated time zones unless the user passes a value they are assigned the default time zone.

```{r}
date_time
date <- dtt_date(date_time)
date
time <- dtt_time(date_time)
time
dtt_date_time(date, time)
```

### Timer

An timer is a hms scalar with a start attribute indicating the time in seconds since 1970-01-01 00:00:00 GMT when it started.
`dtt_elapsed()` returns the duration for which the timer has been
running (see below).

```{r}
timer <- dtt_timer(start = TRUE)
dtt_elapsed(timer)
Sys.sleep(1)
dtt_elapsed(timer)
Sys.sleep(1)
dtt_elapsed(timer)
timer <- dtt_stop(timer)
dtt_elapsed(timer)
Sys.sleep(1)
dtt_elapsed(timer)
timer <- dtt_start(timer)
Sys.sleep(1)
dtt_elapsed(timer)
```

## Inspiration

[lubridate](https://lubridate.tidyverse.org)

## Contribution

Please report any [issues](https://github.com/poissonconsulting/dttr2/issues).

[Pull requests](https://github.com/poissonconsulting/dttr2/pulls) are always welcome.

Please note that the 'dttr2' project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.
